---
title: "QLSC_Assignment_3.1"
author: "Niklas Brake"
date: "October 22, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(DESeq2)

count_matrix = read.table("C:\\Users\\brake\\Documents\\QLSC600\\Module 3\\QLS_counts.tsv");
sample_annotation = read.table("C:\\Users\\brake\\Documents\\QLSC600\\Module 3\\QLS_annotations.tsv");

dds = DESeqDataSetFromMatrix(countData = count_matrix, colData = sample_annotation, design = ~ group)
ddsT <-rlog(dds)
PCs = plotPCA(ddsT, intgroup = "group", ntop = 1000, returnData = TRUE)
PCs

dds = DESeq(dds)
DE_Results = results(dds)
DE_Results[DE_Results$padj < 0.01 & !is.na(DE_Results$padj),]
```

One observes that PC1 segregates the data into group A and group B, with the exception of one A sample, who's PC1 is positive. This is sample A2.

```{r}
count_matrix2 = count_matrix[,c(1:2,4:8)]
sample_annotation2 = sample_annotation[c(1:2,4:8),]
dds2 = DESeqDataSetFromMatrix(countData = count_matrix2, colData = sample_annotation2, design = ~ group)
dds2T <-rlog(dds2)
PCs2 = plotPCA(dds2T, intgroup = "group", ntop = 1000, returnData = TRUE)
PCs2

dds2 = DESeq(dds2)
DE_Results2 = results(dds2)
DE_Results2[DE_Results2$padj < 0.01 & !is.na(DE_Results2$padj),]
```

We now get almsot a four-fold increase in the number of "significant" genes. This makes sense considering we removed the case where a "B-like" sample was labelled A; the two groups are now linearly seperable in the projection onto the first two principle components, whereas before they were not. 

```{r}
ddsshort = dds[order(DE_Results$padj)[c(1:100)],]
heatmap(counts(ddsshort, normalized = TRUE))
```

This presents clustering similar to what was seen in the PCA. However, since we are choosing genes that we already know identify group A from group B, the result of the clustering algorithm is biased to our selection of data fed to it. 

We know that genes with low expression levels tend not the distinguish between groups well. Since this is true regardless of the specific phenotype we are looking at, a less biased sampling would be to select for genes with high expression levels.












